A multidimensional approach for measuring meritocratic beliefs: Advantages, limitations and alternatives to the ISSP social inequality survey

### 2023-10-26 paper-blinded-RR-IJS.Rmd

   **Keywords**: meritocracy, survey methodology, social inequality, construct measurement 
# keywords: "meritocracy, survey methodology, social inequality, construct measurement"
output:
  bookdown::word_document2:
    number_sections: false
    reference_docx: input/template.docx
    toc: false
  bookdown::pdf_document2:
    template: null
    number_sections: false    
    toc: false
    keep_tex: true
    pandoc_args:
     - --template=input/mytemplate.tex #custom template para usar autores con afiliacion
  bookdown::html_document2:
    number_sections: false
linkcolor: blue
bibliography: input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=1in
header-includes:
  - \usepackage{times}           # Times New Roman
#  - \hypersetup {colorlinks = true, linkcolor = blue, urlcolor = blue}    
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage[british,UKenglish,USenglish,american]{babel}
always_allow_html: true # for word output
fontsize: 10pt
papersize: a4
---

@@

  # Declare ggplot2 features
  ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
                                                                             colour = "gray85"),
                                    panel.border = ggplot2::element_blank(),
                                    axis.text.y = ggplot2::element_text(size = 13,
                                                                        hjust = 1),
                                    title = ggplot2::element_text(size = 13,
                                                                  face = "bold"),
                                    legend.text = ggplot2::element_text(size = 12),
                                    plot.caption = ggplot2::element_text(size = 10,
                                                                         face = "plain",
                                                                         hjust = 1)))
  #Plot 1: Perception meritocratic
  p1<- likert_issp %>%
    dplyr::select(hwork_r:ambition_r) %>%
    sjPlot::plot_likert(geom.colors = "PuBu",
                        grid.range  =  c (1.2 , 1.2),
                        title = "a. Perception meritocratic",
                          geom.size = c(0.62),
                          catcount = 5,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = F,
                          show.n = FALSE) +
    ggplot2::theme(legend.position = "none")
  #Plot 2: Perception Non-meritocratic
  p2<-likert_issp %>%
    dplyr::select(wealthy_r:polcone_r) %>%
    sjPlot::plot_likert(geom.colors = "PuBu",
                        grid.range  =  c (1.2 , 1.2),
                        title = "b. Perception Non-meritocratic",
                          geom.size = c(0.9),
                          catcount = 5,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = F,
                          show.n = FALSE)  +
    ggplot2::theme(legend.position = "none")
  #Plot 3: Preference meritocratic
  p3<-likert_issp %>%
    dplyr::select(welljob_r:hardjob_r) %>%
    sjPlot::plot_likert(geom.colors = "PuBu",
                        grid.range  =  c (1.2 , 1.2),
                        title = "c. Preference meritocratic",
                          geom.size = c(0.62),
                          catcount = 5,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = F,
                          show.n = FALSE) +
    ggplot2::theme(legend.position = "none")
  #Plot 4: Preference Non-meritocratic
  p4<-likert_issp %>%
    dplyr::select(family_r:child_r) %>%
    sjPlot::plot_likert(geom.colors = "PuBu",
                        grid.range  =  c (1.2 , 1.2),
                        title = "d. Preference Non-meritocratic",
                          geom.size = c(0.62),
                          catcount = 5,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = F,
                          show.n = FALSE) +
    ggplot2::theme(legend.position = "none")

  #Legend: Extract legend for the final plot using ggpubr
  leg<- ggpubr::as_ggplot(
    ggpubr::get_legend(
      dplyr::select(likert_issp, family_r:child_r) %>%
        sjPlot::plot_likert(geom.colors = "PuBu", catcount = 5) +
        ggplot2::scale_fill_brewer(
          breaks = c(5:1),
          direction = -1,
          labels = c("Not important at all","Not very important","Fairly important",
                     "Very important","Essential")) +
        ggplot2::theme(legend.position = "bottom")
      )
    )
```

@@

  #Plot the matrix using corrplot
  corrplot::corrplot(cor_issp,
    method = "color",
    addCoef.col = "#000390",
    type = "upper",
    tl.col = "black",
    col=colorRampPalette(c("white","#0068DC"))(12),
    bg = "white",
    na.label = "-")
    # dev.off()
```

@@

# create figure 1: perception merit by preferences merit
figure1 <- 
ggplot(avg_count, aes(x=percmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(avg_count$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+
  geom_smooth(method = "lm",fullrange=TRUE) +
  xlab("Meritocratic perception")+
  ylab("Meritocratic preference")  
# create figure 2: perception non-merit by preferences merit
figure2 <- 
ggplot(avg_count, aes(x=percnmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(avg_count$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+ 
  geom_smooth(method = "lm",fullrange=TRUE) +  
  geom_smooth(method = "lm") +
  xlab("Non-meritocratic perception")+
  ylab("Meritocratic preference")

@@

# Western countries-------------------------------------------------------------
wesmerit <- 
indus %>%  
ggplot(aes(x=percmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(indus$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+
  geom_smooth(method = "lm",fullrange=TRUE) +
  xlab("Meritocratic perception")+
  ylab("Meritocratic preference") +
  ggtitle("Industrialized countries")
wesnonmer <- 
indus %>%  
ggplot(aes(x=percnmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(indus$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+ 
  geom_smooth(method = "lm",fullrange=TRUE) +
  xlab("Non-meritocratic perception")+
  ylab("Meritocratic preference") +
  ggtitle("Industrialized countries")

# Non western countries --------------------------------------------------------
nonwesmer <- 
lateindus %>%  
ggplot(aes(x=percmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(lateindus$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+
  geom_smooth(method = "lm",fullrange=TRUE) +
  xlab("Meritocratic perception")+
  ylab("Meritocratic preference") +
  ggtitle("Late industralized countries")
nonwesnonmer <- 
lateindus %>%  
ggplot(aes(x=percnmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(lateindus$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+
  geom_smooth(method = "lm",fullrange=TRUE) +
  xlab("Non-meritocratic perception")+
  ylab("Meritocratic preference") +
  ggtitle("Late industralized countries")

@@

  #Plot 1: Perception
  p1<-
    dplyr::select(dat_likert_s2,perc_effort_r:perc_netw_r) %>%
    sjPlot::plot_likert(geom.colors = "PuBu",
                        title = c("a. Perceptions"),
                        geom.size = 0.8,
                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
                        catcount = 4,
                        cat.neutral = 3,
                        grid.range  =  c (1.2 , 1.2),
                        values  =  "sum.outside",
                        reverse.colors = T,
                        reverse.scale = F,
                        show.n = FALSE) +
    ggplot2::theme(legend.position = "none")

  #Plot 2: Preferences
  p2<-
    dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
    sjPlot::plot_likert(geom.colors = "PuBu",
                        title = c("b. Preferences"),
                        geom.size = 0.8,
                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
                        catcount = 4,
                        cat.neutral = 3,
                        grid.range  =  c (1.2 , 1.2),
                        values  =  "sum.outside",
                        reverse.colors = T,
                        reverse.scale = F,
                        show.n = FALSE)  +
    ggplot2::theme(legend.position = "none")

  #Legend: Extract legend for the final plot using ggpubr
  leg<- ggpubr::as_ggplot(
    ggpubr::get_legend(
    dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
        sjPlot::plot_likert(geom.colors = "PuBu",
                            catcount = 4,
                            cat.neutral = 3) +
      ggplot2::theme(legend.position = "bottom") +
      ggplot2::scale_fill_manual(
        values = c("#b3b3b3ff", "#f1eef6ff", "#bdc9e1ff","#74a9cfff","#0570b0ff"),
        breaks=c("neutral",4,3,2,1),
        labels=c("Neither agree nor disagree","Strongly disagree","Disagree",
                 "Agree","Totally agree")
        )
      )
    )
```

@@

  corrplot::corrplot(cor_s2,
    method = "color",
    addCoef.col = "#000390",
    type = "upper",
    tl.col = "black",
    col=colorRampPalette(c("white","#0068DC"))(8),
    bg = "white",
    na.label = "-")
  # dev.off()
```

>>>>

   **Keywords**: meritocracy, survey methodology, social inequality, construct measurement 
# keywords: "meritocracy, survey methodology, social inequality, construct measurement"
output:
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    template: null
    number_sections: false    
    toc: false
    keep_tex: true
    pandoc_args:
     - --template=input/mytemplate.tex #custom template para usar autores con afiliacion
  bookdown::word_document2:
    number_sections: false
    reference_docx: input/template.docx
    toc: false
linkcolor: blue
bibliography: input/bib/meritocracy.bib
csl: input/bib/apa7-singlespace.csl
editor_options:
  chunk_output_type: console
geometry: margin=1in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage[british,UKenglish,USenglish,american]{babel}
always_allow_html: true # for word output
fontsize: 10pt
papersize: a4
---

@@

  # Declare ggplot2 features
  ggplot2::theme_set(ggplot2::theme(panel.background = ggplot2::element_rect(fill = "gray85",
                                                                             colour = "gray85"),
                                    panel.border = ggplot2::element_blank(),
                                    axis.text.y = ggplot2::element_text(size = 13,
                                                                        hjust = 1),
                                    title = ggplot2::element_text(size = 13,
                                                                  face = "bold"),
                                    legend.text = ggplot2::element_text(size = 12),
                                    plot.caption = ggplot2::element_text(size = 10,
                                                                         face = "plain",
                                                                         hjust = 1)))
  #Plot 1: Perception meritocratic
  p1<- likert_issp %>%
    dplyr::select(hwork_r:ambition_r) %>%
    sjPlot::plot_likert(geom.colors = "Greys",
                        grid.range  =  c (1.2 , 1.2),
                        title = "a. Perception meritocratic",
                          geom.size = c(0.62),
                          catcount = 5,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = F,
                          show.n = FALSE) +
    ggplot2::theme(legend.position = "none")
  #Plot 2: Perception Non-meritocratic
  p2<-likert_issp %>%
    dplyr::select(wealthy_r:polcone_r) %>%
    sjPlot::plot_likert(geom.colors = "Greys",
                        grid.range  =  c (1.2 , 1.2),
                        title = "b. Perception Non-meritocratic",
                          geom.size = c(0.9),
                          catcount = 5,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = F,
                          show.n = FALSE)  +
    ggplot2::theme(legend.position = "none")
  #Plot 3: Preference meritocratic
  p3<-likert_issp %>%
    dplyr::select(welljob_r:hardjob_r) %>%
    sjPlot::plot_likert(geom.colors = "Greys",
                        grid.range  =  c (1.2 , 1.2),
                        title = "c. Preference meritocratic",
                          geom.size = c(0.62),
                          catcount = 5,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = F,
                          show.n = FALSE) +
    ggplot2::theme(legend.position = "none")
  #Plot 4: Preference Non-meritocratic
  p4<-likert_issp %>%
    dplyr::select(family_r:child_r) %>%
    sjPlot::plot_likert(geom.colors = "Greys",
                        grid.range  =  c (1.2 , 1.2),
                        title = "d. Preference Non-meritocratic",
                          geom.size = c(0.62),
                          catcount = 5,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = F,
                          show.n = FALSE) +
    ggplot2::theme(legend.position = "none")

  #Legend: Extract legend for the final plot using ggpubr
  leg<- ggpubr::as_ggplot(
    ggpubr::get_legend(
      dplyr::select(likert_issp, family_r:child_r) %>%
        sjPlot::plot_likert(geom.colors = "Greys", catcount = 5) +
        ggplot2::scale_fill_brewer(
          breaks = c(5:1),
          direction = -1,
          labels = c("Not important at all","Not very important","Fairly important",
                     "Very important","Essential"),palette = "Greys") +
        ggplot2::theme(legend.position = "bottom")
      )
    )
```

@@

  #Plot the matrix using corrplot
  corrplot::corrplot(cor_issp,
    method = "color",
    # addCoef.col = "#000390",
    addCoef.col = "Gray",
    type = "upper",
    tl.col = "black",
    # col=colorRampPalette(c("white","#0068DC"))(12),
    col=colorRampPalette(c("white","gray9"))(12),
    bg = "white",
    na.label = "-")
    # dev.off()
```

@@

# create figure 1: perception merit by preferences merit
figure1 <- 
ggplot(avg_count, aes(x=percmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(avg_count$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+
  geom_smooth(method = "lm",fullrange=TRUE,color="black") +
  xlab("Meritocratic perception")+
  ylab("Meritocratic preference")  
# create figure 2: perception non-merit by preferences merit
figure2 <- 
ggplot(avg_count, aes(x=percnmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(avg_count$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+ 
  geom_smooth(method = "lm",fullrange=TRUE,color="black") +  
  xlab("Non-meritocratic perception")+
  ylab("Meritocratic preference")

@@

# Western countries-------------------------------------------------------------
wesmerit <- 
indus %>%  
ggplot(aes(x=percmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(indus$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+
  # geom_smooth(method = "lm",fullrange=TRUE) +
geom_smooth(method = "lm",fullrange=TRUE,color="black") +  
  xlab("Meritocratic perception")+
  ylab("Meritocratic preference") +
  ggtitle("Industrialized countries")
wesnonmer <- 
indus %>%  
ggplot(aes(x=percnmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(indus$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+ 
  # geom_smooth(method = "lm",fullrange=TRUE) +
geom_smooth(method = "lm",fullrange=TRUE,color="black") +    
  xlab("Non-meritocratic perception")+
  ylab("Meritocratic preference") +
  ggtitle("Industrialized countries")

# Non western countries --------------------------------------------------------
nonwesmer <- 
lateindus %>%  
ggplot(aes(x=percmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(lateindus$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+
  # geom_smooth(method = "lm",fullrange=TRUE) +
  geom_smooth(method = "lm",fullrange=TRUE,color="black") +  
  xlab("Meritocratic perception")+
  ylab("Meritocratic preference") +
  ggtitle("Late industralized countries")
nonwesnonmer <- 
lateindus %>%  
ggplot(aes(x=percnmerit, y=prefmerit)) +
  # geom_point() + 
  geom_text(label=as.character(lateindus$clab)) +
  scale_x_continuous(limits = c(-0.6,0.6))+
  scale_y_continuous(limits = c(-0.6,0.6))+
  geom_smooth(method = "lm",fullrange=TRUE) +
  geom_smooth(method = "lm",fullrange=TRUE,color="black") +    
  xlab("Non-meritocratic perception")+
  ylab("Meritocratic preference") +
  ggtitle("Late industralized countries")

@@

  #Plot 1: Perception
  p1<-
    dplyr::select(dat_likert_s2,perc_effort_r:perc_netw_r) %>%
    sjPlot::plot_likert(geom.colors = "Greys",
                        title = c("a. Perceptions"),
                        geom.size = 0.8,
                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
                        catcount = 4,
                        cat.neutral = 3,
                        grid.range  =  c (1.2 , 1.2),
                        values  =  "sum.outside",
                        reverse.colors = T,
                        reverse.scale = F,
                        show.n = FALSE) +
    ggplot2::theme(legend.position = "none")

  #Plot 2: Preferences
  p2<-
    dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
    sjPlot::plot_likert(geom.colors = "Greys",
                        title = c("b. Preferences"),
                        geom.size = 0.8,
                        axis.labels = c("Effort", "Talent", "Rich Parents", "Contacts"),
                        catcount = 4,
                        cat.neutral = 3,
                        grid.range  =  c (1.2 , 1.2),
                        values  =  "sum.outside",
                        reverse.colors = T,
                        reverse.scale = F,
                        show.n = FALSE)  +
    ggplot2::theme(legend.position = "none")

  #Legend: Extract legend for the final plot using ggpubr
  leg<- ggpubr::as_ggplot(
    ggpubr::get_legend(
    dplyr::select(dat_likert_s2,pref_effort_r:pref_netw_r) %>%
        sjPlot::plot_likert(geom.colors = "Greys",
                            catcount = 4,
                            cat.neutral = 3) +
      ggplot2::theme(legend.position = "bottom") +
      ggplot2::scale_fill_manual(
        values = c("#B3B3B3","#F7F7F7","#CCCCCC","#969696","#525252"),
        # values = c("#b3b3b3ff", "#f1eef6ff", "#bdc9e1ff","#74a9cfff","#0570b0ff"),
        breaks=c("neutral",4,3,2,1),
        labels=c("Neither agree nor disagree","Strongly disagree","Disagree",
                 "Agree","Totally agree")
        )
      )
    )
```

@@

  corrplot::corrplot(cor_s2,
    method = "color",
    # addCoef.col = "#000390",
    addCoef.col = "Gray",
    type = "upper",
    tl.col = "black",
    # col=colorRampPalette(c("white","#0068DC"))(8),
    col=colorRampPalette(c("white","gray9"))(12),
    bg = "white",
    na.label = "-")
  # dev.off()
```
